---
title: "Power Enhancement for High-Dimensional Mediation: A Guided Tour"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Power Enhancement for High-Dimensional Mediation: A Guided Tour}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.2)
library(POEMED)
# Rejection rates quoted in the prose, as percentages. With 40 replications
# a rate such as 0.225 falls on a half percent (22.5), so the half is kept
# instead of being rounded to a whole number in one direction or the other.
pct <- function(rate) format(100 * rate, digits = 3)
```

This vignette is written for a reader who is comfortable with regression
but new to high-dimensional mediation. It explains the problem the
package solves, builds up the idea behind the power-enhanced tests,
then walks through a first analysis, the output, the different outcome
types, a simulation study, and finally the real-data application from
the article that introduced the tests, Yu and Kelley (in press). You do
not need to have read the article first; pointers to it appear where the
details live.

## 1. What Is a Mediation Analysis?

A mediation analysis asks *how* an exposure affects an outcome. Instead
of only asking whether a treatment `X` changes an outcome `Y`, it asks
whether `X` works **through** an intermediate variable `M`, the
*mediator*. The classic picture is a chain:

```
        a                 b
   X  ----->   M   ----->   Y
    \                       /
     \________ c' _________/
            (direct)
```

The effect of `X` on `Y` that travels through `M` is the **indirect
effect**: in a linear model it is the product `a * b` of the
`X -> M` path and the `M -> Y` path. The part of `X`'s effect that does
not go through `M` is the **direct effect** `c'`. Their sum is the
**total effect**.

In modern applications there is rarely a single mediator. Genomics
studies screen thousands of methylation sites; the application later in
this vignette has 57 health-expenditure indicators. When there are many
candidate mediators `M = (M_1, ..., M_p)`, the natural first question is
simply:

> **Is there any active mediator at all?** That is, does `X` affect `Y`
> through *any* of the candidate mediators?

This is a single global hypothesis test, and it is what POEMED is built
to perform with high power.

## 2. The Problem: Total-Indirect-Effect Tests Have a Blind Spot

With many mediators, the conventional approach summarizes them by the
**total indirect effect**, written `beta`: the sum over all mediators of
each one's individual indirect effect. The global test then checks
whether `beta = 0`.

This works well when the mediators all push the outcome the same way.
But consider two active mediators whose individual indirect effects are
equal in size and **opposite in sign**: say `+0.5` through `M_1` and
`-0.5` through `M_2`. Both mediators are genuinely active, yet their
contributions cancel, so the total indirect effect is

```
beta = (+0.5) + (-0.5) = 0.
```

A test built on `beta` sees zero and concludes "no mediation," even
though two mediators are plainly active. This is not a rare edge case:
whenever different pathways move an outcome in opposite directions
(**heterogeneous** mediation, and in the extreme **contrasting**
mediation), the total indirect effect understates the truth and the test
loses power. The article calls this out as the core limitation of
existing high-dimensional mediation tests.

## 3. The Idea: A Power Enhancement Component

The power-enhanced (PE) test fixes the blind spot by adding a second
piece that cannot cancel. After a penalized fit narrows the candidates to
a small set of mediators, the PE component `J_m` adds up, over those
mediators, the **magnitude** of each one's combined path signal. In code
notation, with `alpha_hat[j]` the estimated `M -> Y` coefficient of
mediator `j`, `Gamma_hat[i, j]` the estimated `X -> M` coefficient from
exposure `i`, `se_*` their standard errors, and `p_alpha[j]` and
`p_Gamma[i, j]` the two paths' p-values:

```
J_m = sqrt(p) * sum over exposures i and selected mediators j of
      |alpha_hat[j] / se_alpha[j]| * |Gamma_hat[i, j] / se_Gamma[i, j]|
      * 1{max(p_alpha[j], p_Gamma[i, j])
            <= error_level / (s * q * log(log(n)))}
```

Here `s` is the number of mediators the penalized fit selected, `q` the
number of exposures, and `n` the sample size. This is the default
Bonferroni screen; the `"BH"` and `"BY"` methods of Section 6 compare
adjusted p-values with `error_level / log(log(n))` instead.

Two things matter here. First, the indicator keeps a mediator only when
**both** its `X -> M` path and its `M -> Y` path pass that strict,
multiplicity-adjusted threshold, so noise mediators do not contribute.
The threshold is set by `error_level` (default 0.05), the error rate
targeted for the identified mediators, so it shapes the global test as
well (Section 6 shows how). Second, because `J_m` sums **absolute
values**, a `+0.5` mediator and a `-0.5` mediator *reinforce* each other
instead of canceling.

The full test statistic is `M_PE = S_n + J_m`, the benchmark Wald
statistic `S_n` (the total-indirect-effect test) plus the enhancement.
Under the global null of no mediation, the screen keeps no mediator with
probability tending to one as `n` grows, so `J_m = 0` and `M_PE` has the
same chi-square reference distribution as `S_n` asymptotically. In a
finite sample `J_m` is occasionally positive under the null, and because
`M_PE` is never smaller than `S_n`, the PE test then rejects a little more
often than the benchmark (the simulation in Section 8 shows this). Under
heterogeneous alternatives, `J_m` grows large and the test gains power.
That is the whole idea: **keep the size in large samples, and add power
where the old test is blind.**

The same indicator that builds `J_m` also tells you *which* mediators
are active, targeting the familywise error rate (the default) or the
false discovery rate. The tuning grid of the penalized fit shapes both
the global test and the selected set; Sections 4, 8, and 11 say how to
read it and how to tune it.

## 4. A First Analysis, Step by Step

The package can simulate data so you can see the method work before
bringing your own. `simulate_mediation_data()` generates the patterns
studied in the article. Here is a **contrasting** setting: a few active
mediators whose effects cancel to a zero total indirect effect.

```{r sim}
set.seed(113)
d <- simulate_mediation_data(n = 200, p = 60, outcome = "continuous",
                             pattern = "contrasting", c1 = 1)
```

The result is a list. The four pieces an analysis needs are the
exposure, outcome, mediators, and (optionally) confounders:

```{r sim-pieces}
dim(d$X)   # exposure: 200 observations, 1 column
length(d$Y)   # outcome: 200 values
dim(d$M)   # mediators: 200 x 60
d$active_mediators   # which mediators are truly active (the ground truth)
d$beta               # the total indirect effect: zero, by construction
```

So four mediators are genuinely active, but the total indirect effect is
zero because their effects cancel. This is precisely the situation that
defeats a total-indirect-effect test. Now run the global test:

```{r first-fit}
fit <- pe_mediation(d$X, d$Y, d$M, outcome = "continuous")
fit
```

Read the two p-values. `pval_hdmm` is the benchmark Wald test on the
total indirect effect; it is large, so that test fails to detect any
mediation (exactly the blind spot from Section 2). `pval_pe` is the
power-enhanced test; it is essentially zero, so it detects the mediation
decisively. The footer lists the mediators the screen selected, here
three of the four truly active mediators. The power enhancement, made
concrete.

The last footer line reports the tuning parameter `lambda` that the
high-dimensional BIC (HBIC) chose for the penalized fit and the grid it
searched: by default 100 values from 0.05 to 10, the same grid for every
outcome type. HBIC is minimized over the grid alone, so whatever grid is
offered, the value chosen is the best of the values offered. Here the
smallest value won. The footer says so, and the fit warned, because a
minimum at an end of the grid may lie beyond that end or between it and
its neighbor; a grid that extends further, or is spaced more finely, is
worth a look whenever that happens. Doing that here moves the chosen
value inside the grid and leaves the selected mediators unchanged:

```{r finer-grid}
finer <- pe_mediation(d$X, d$Y, d$M, outcome = "continuous",
                      lambda_grid = seq(0.01, 10, length.out = 200))
attr(finer, "tuning")$lambda_selected
attr(finer, "selected_mediators")
```

The grid can be tuned the other way as well: fewer values, or a range
narrowed to where the HBIC minimum has been seen to fall, make each fit
faster, which matters when many fits are run (Section 8 does this for a
simulation study). The chunks below fit on the default grid, and each
would repeat the warning above; it is not shown again.

## 5. Reading the Result Table

Every POEMED test returns a tidy table that prints with sensible rounding
(whole numbers without decimals, p-values to four places) while storing
full precision underneath. The rows are:

| term | meaning |
|---|---|
| `stat_hdmm`, `pval_hdmm` | the benchmark Wald statistic `S_n` and its p-value (the total-indirect-effect test of Guo et al.) |
| `stat_pe`, `pval_pe` | the power-enhanced statistic `M_PE` and its p-value |
| `j_pe` | the power enhancement component `J_m` (the amount added to `S_n`) |
| `total_indirect_effect` | the estimated total indirect effect `beta` |
| `n_selected_mediators` | how many individual mediators the screen selected |
| `df` | degrees of freedom of the chi-square reference (the number of exposures) |
| `n_candidate_mediators`, `n_observations` | `p` and `n` |

The table gives a point estimate of `beta` but no confidence interval,
and the article reports none. In simulations, a Wald interval built on
the penalized estimate fell short of its nominal coverage, most clearly
for binary and count outcomes, so the package does not report one.

To pull a number out, index the table like any data frame; the stored
value keeps full precision even though the display rounds:

```{r read}
fit$value[fit$term == "pval_pe"]            # the exact p-value
attr(fit, "selected_mediators")             # the selected mediator indices
```

## 6. Identifying Which Mediators Are Active

The selected set is controlled by the `method` argument. The default,
`"Bonferroni"`, targets the familywise error rate and is the most
conservative of the three, so it may miss weak mediators. Its control of
that rate is asymptotic: in finite samples the rate also depends on the
design and on the tuning grid, since a grid that reaches small values of
`lambda` lets more candidates into the penalized fit. The `"BH"` and `"BY"` options control the false discovery rate
instead and typically recover more mediators.

```{r methods, warning = FALSE}
des_idx <- function(f) attr(f, "selected_mediators")
des_idx(pe_mediation(d$X, d$Y, d$M, outcome = "continuous", method = "Bonferroni"))
des_idx(pe_mediation(d$X, d$Y, d$M, outcome = "continuous", method = "BH"))
```

The target error rate is `error_level` (default 0.05): the familywise
error rate under `"Bonferroni"`, the false discovery rate under `"BH"`
and `"BY"`. It is distinct from the significance level you use to read
`pval_pe`, but the global test depends on it as well, because it sets
the screening threshold inside `J_m` (Section 3). A stricter value can
therefore remove the enhancement on a given data set:

```{r error-level, warning = FALSE}
strict <- pe_mediation(d$X, d$Y, d$M, outcome = "continuous",
                       error_level = 0.005)
strict[strict$term %in% c("pval_hdmm", "j_pe", "pval_pe"), ]
```

On these data a Bonferroni threshold built from `error_level = 0.005`
keeps no mediator, so `J_m` is zero and `pval_pe` equals the benchmark's
`pval_hdmm`, where the default 0.05 gave a p-value of essentially zero.

## 7. Binary and Count Outcomes

The same construction works for non-continuous outcomes through a link
function. Name the outcome type and the front end dispatches to the
logistic (binary) or Poisson (count) model. The total indirect effect is
then on the link scale (log-odds for binary, log-mean for count).

```{r glm, warning = FALSE}
set.seed(113)
db <- simulate_mediation_data(n = 250, p = 50, outcome = "binary",
                              pattern = "contrasting", c1 = 1, c2 = 1)
pe_mediation(db$X, db$Y, db$M, outcome = "binary")
```

The tuning grid is the same for every outcome type, 100 values from
0.05 to 10, and is tuned the same way (Section 4).

## 8. A Simulation Study: Size and Power

`pe_power_curve()` runs Monte Carlo studies of the kind the article
reports. It
sweeps the signal strength `c1`, simulating many data sets at each value
and recording how often each test rejects. At `c1 = 0` (no mediation)
the rejection rate estimates the Type I error rate; at nonzero `c1` it
estimates power. A rate from `n_rep` replications carries a Monte Carlo
standard error of about `sqrt(r * (1 - r) / n_rep)`, so the 40-replication
run below knows a rate near 0.05 to within about 0.03 and a rate near 0.5
to within about 0.08.

A simulation study is also where the tuning grid is worth tuning to the
situation. Each replication searches the whole grid, so the default grid
of 100 values costs about five times the fitting of a 20-value grid. The
study below passes a shorter grid, 20 values from 0.05 to 2: the fit of
Section 4 put the HBIC minimum at the bottom of the default grid, so the
shorter grid keeps that end and drops the range above 2, which no fit at
this design reaches. Whatever the grid, each fit keeps the best value the
grid offers, and the warning a single fit gives when its choice sits at
an end of the grid is not raised inside a study, which searches a fixed
grid by design. (On these 120 simulated data sets the default grid gives
the same six rejection rates; it only takes five times as long.)

```{r power, fig.alt = "Rejection-rate curves for the benchmark and power-enhanced tests"}
set.seed(113)
pc <- pe_power_curve(n = 150, p = 50, outcome = "continuous",
                     pattern = "contrasting", c1_grid = c(0, 0.5, 1),
                     n_rep = 40, lambda_grid = seq(0.05, 2, length.out = 20))
pc

plot(pc, main = "Contrasting Mediation: PE vs. Benchmark")
```

At `c1 = 0` the benchmark rejects in `r pct(pc$rejection_hdmm[pc$c1 == 0])`
percent of replications and the power-enhanced test in
`r pct(pc$rejection_pe[pc$c1 == 0])` percent, both near the nominal 5 percent
given the Monte Carlo error of 40 replications. The PE rate can never
fall below the benchmark's on the same data sets, because `M_PE` is never
smaller than `S_n`. It is higher whenever a null replication has a
positive `J_m` that carries `M_PE` past the critical value where `S_n`
alone falls short (Section 3). As the contrasting signal grows the
benchmark rejects in `r pct(pc$rejection_hdmm[pc$c1 == 1])` percent of
replications at `c1 = 1` and the power-enhanced test in
`r pct(pc$rejection_pe[pc$c1 == 1])` percent. That is the enhancement at work.
The article's own study, at `n = 300`, `p = 500`, and 1,000 replications
per point, is the converged version of this picture; the reproduction
vignette gives the calls, with the grids of the article's scripts.

## 9. The Real Application: Health Spending as a Mediator

The package ships the article's empirical data, `WHO_health_mediation`:
a panel of 91 WHO member states over 2000--2021. The exposure is annual
growth in GDP per capita, the candidate mediators are 57
health-expenditure indicators from the WHO Global Health Expenditure
Database, and there are five health outcomes (infant and under-five
mortality, life expectancy, low birthweight, and undernourishment). The
substantive question: **does health spending mediate the relationship
between economic growth and population health?**

```{r who-data}
data(WHO_health_mediation)
dim(WHO_health_mediation)
WHO_health_mediation[1:3, c("country", "region", "income", "year",
                            "gdp_growth", "imr", "gge_gdp")]
```

The 57 mediators are strongly intercorrelated (many are different
normalizations of the same spending), which is exactly the
high-dimensional, correlated-mediator regime the method targets. Their
definitions are in `WHO_indicator_codebook`:

```{r codebook}
head(WHO_indicator_codebook, 4)
```

### Building One Analysis Design

`WHO_mediation_design()` assembles the exposure, outcome, mediator, and
confounder matrices for a chosen outcome, handling the missing-data
coverage and the confounders (region, income group, and year) for you.
The call below follows the preprocessing of the authors' analysis code
behind the article's tables: it standardizes the exposure and the
mediators, centers the outcome, and leaves the confounders (the region
and income indicators and the year) unscaled, then passes `scale = FALSE`
so that `pe_mediation()` fits the matrices as given. Its tuning grid,
`seq(0.1, 2, length.out = 100)`, is the grid behind the article's
real-data tables and the default of `WHO_mediation_analysis()` for the
global and every subgroup model. In this global model HBIC picks 0.1,
the smallest value of that grid, as it does in about half of the
article's cells; the article's analysis keeps its grid as it is, and so
does this vignette, so the warning a single fit gives at an end of its
grid is not shown in this section. The article's text says
that covariates are standardized before fitting, but its global finding
for infant mortality is recovered with the confounders left unscaled.
The last part of this section shows what standardizing them as well
changes.

```{r who-design, warning = FALSE}
des <- WHO_mediation_design("imr")          # infant mortality rate
c(observations = des$n, mediators = ncol(des$M), confounders = ncol(des$Z))

fit <- pe_mediation(scale(des$X), des$Y - mean(des$Y), scale(des$M),
                    Z = des$Z, outcome = "continuous", scale = FALSE,
                    lambda_grid = seq(0.1, 2, length.out = 100))
selected <- des$indicators[attr(fit, "selected_mediators")]
WHO_indicator_codebook$description[
  match(selected, WHO_indicator_codebook$indicator)]
```

The PE test identifies general government expenditure as a share of GDP
(`gge_gdp`) as the mediator through which economic growth reaches infant
mortality. Yu and Kelley (in press) report the same single indicator for
the global model in their Table 1.

### The Full Subgroup Analysis

`WHO_mediation_analysis()` runs the whole set of models the article
reports for one outcome (global, then within each WHO region, income
group, and region-income cell), prepares the data the same way as the
fit above, and returns one row per group, laid out like the article's
Table 1 (infant mortality) and Table 2 (undernourishment). Here it fits
the global model and the six regions on its default tuning grid.

```{r who-analysis}
res <- WHO_mediation_analysis("imr", groupings = c("global", "region"))
res[, c("group", "n_countries", "pval_hdmm", "pval_pe", "selected_mediators")]
```

Compare the benchmark and PE columns. Where the benchmark test sees
nothing (a large `pval_hdmm`) while the PE test rejects decisively and
names the indicators responsible, heterogeneous mediation is hiding the
signal from the conventional test. In the Western Pacific region, for
example, the benchmark p-value is
`r format(res$pval_hdmm[res$group == "WPR"], digits = 2)`, yet the PE
test rejects and flags compulsory health insurance as a share of current
health expenditure (`chi_che`). Table 1 of Yu and Kelley (in press)
shows the same contrast for this region.

`summary()` reads that comparison off the table for you, counting the
groups where the PE test detects mediation the benchmark misses:

```{r who-summary}
summary(res)
```

### Why the Preprocessing Matters

Passing the raw design to `pe_mediation()` with the default
`scale = TRUE` standardizes the confounders as well, and that is a
different model. On the same tuning grid as the first fit:

```{r who-raw, warning = FALSE}
raw <- pe_mediation(des$X, des$Y, des$M, Z = des$Z, outcome = "continuous",
                    lambda_grid = seq(0.1, 2, length.out = 100))
c(confounders_unscaled = fit$value[fit$term == "pval_hdmm"],
  all_standardized = raw$value[raw$term == "pval_hdmm"])
des$indicators[attr(raw, "selected_mediators")]
```

The benchmark Wald p-value moves from about
`r format(fit$value[fit$term == "pval_hdmm"], digits = 2)` to about
`r format(raw$value[raw$term == "pval_hdmm"], digits = 2)`, and the
screen selects a second indicator, social health insurance
(`shi_che`). Leave the confounders unscaled (or call
`WHO_mediation_analysis()`, which does) whenever the goal is to compare
results with the article's tables.

## 10. Beyond the Global Test: The Toolkit

The global test is the core, but POEMED ships a few conveniences around it.

### A Formula Interface for Data Frames

If your variables live in a data frame, `pe_mediate()` takes a formula
for the outcome and exposure, the names of the mediator columns, and
(optionally) confounder names, and calls `pe_mediation()` for you.

```{r formula, warning = FALSE}
set.seed(113)
d <- simulate_mediation_data(n = 200, p = 60, outcome = "continuous",
                             pattern = "contrasting", c1 = 1)
df <- data.frame(y = d$Y, x = d$X[, 1], d$M)
meds <- grep("^X", names(df), value = TRUE)   # the mediator columns
fit <- pe_mediate(y ~ x, data = df, mediators = meds, outcome = "continuous")
```

### Why Each Mediator? The Per-Mediator Table

`pe_mediators()` opens up the screening step: for every candidate
mediator the penalized fit kept, it shows the two path statistics, the
screening p-value (the larger of the two path p-values), and whether the
screen selected the mediator (the `selected` column). This is how you
see *why* the test reached its conclusion.

```{r mediators}
pe_mediators(fit)
```

`plot()` shows the same evidence graphically, the screening strength of
each candidate the penalized fit kept, with the mediators the screen
selected highlighted and the rest marked as screened out.

```{r plot-mediators, fig.alt = "Per-mediator screening evidence for the candidate mediators the penalized fit kept"}
plot(fit)
```

### Working With broom

`tidy()` returns the per-mediator table and `glance()` a one-row summary,
so a POEMED result drops into broom and tidyverse workflows.

```{r broom}
glance(fit)
```

### All Three Multiplicity Methods at Once

Pass `report_all_methods = TRUE` and `pe_selection()` reports the selected
set under Bonferroni (familywise error rate), Benjamini-Hochberg, and
Benjamini-Yekutieli (false discovery rate) side by side, as the
supplement of the article does.

```{r selection, warning = FALSE}
fit_all <- pe_mediation(d$X, d$Y, d$M, outcome = "continuous",
                        report_all_methods = TRUE)
pe_selection(fit_all)
```

### Planning the Sample Size

`ss_power_pe_mediation()` plans a study: over a grid of candidate sample
sizes it estimates the power of the PE test by simulation and stores the
smallest sample size that reaches the target power in the
`"recommended_n"` attribute. That attribute is `NA` when no size in the
grid reaches the target, so choose a grid that brackets it. Under the
default tuning grid this contrasting design reaches a power of 0.8
between 100 and 150 observations, so the grid below brackets that. It
runs a Monte Carlo study (400 fits on the default grid, about a minute
and a half), so it is shown here but not evaluated; a shorter tuning
grid, passed as `lambda_grid`, would cut that time.

```{r ssplan, eval = FALSE}
plan <- ss_power_pe_mediation(outcome = "continuous", pattern = "contrasting",
                              p = 50, n_grid = c(100, 150, 200, 250),
                              target_power = 0.8, n_rep = 100, seed = 113)
plan
attr(plan, "recommended_n")
```

Run as shown, it estimates the power of the power-enhanced test at 0.77,
0.92, 1.00, and 1.00 for the four sizes and recommends `n = 150`. Each
power estimate from 100 replications carries a Monte Carlo standard error
of about 0.04 near a power of 0.8, so a size whose estimate sits just
above the target could give way to the next size under another seed;
raise `n_rep` for a plan you will report.

## 11. Bringing Your Own Data

To analyze your own study, arrange:

- `X`: an `n`-by-`q` matrix of exposures (often one column).
- `Y`: a length-`n` outcome (continuous, 0/1, or counts).
- `M`: an `n`-by-`p` matrix of candidate mediators (`p` may exceed `n`).
- `Z` (optional): an `n`-by-`d` matrix of confounders.

then call `pe_mediation(X, Y, M, Z, outcome = ...)`. The defaults
standardize the inputs as the method assumes, choose the tuning
parameter by a high-dimensional information criterion over a grid of
100 values from 0.05 to 10 for every outcome type, and target the
familywise error rate when selecting individual mediators. The grid can
be tuned: shortened or narrowed to save time, widened or refined when
the value chosen sits at an end of the grid, which the fit warns about
and the footer notes (Section 4). Everything returned is an ordinary
data frame you can index, save, or plot, and the print footer reports
the tuning parameter chosen and the grid it came from.

### When POEMED Is the Right Tool, and When It Is Not

The tests assume a linear (or generalized linear) outcome model with a
sparse set of active mediators, standardized inputs, and complete data.
The section "When to Use POEMED" of `?pe_mediation` states the practical
limits seen on simulated data, with the design behind each. In brief:

- Sample size. The penalized selection needs `n` large relative to
  `log(p)` and to the signal. With `n = 100` and `p = 2000` under the
  contrasting pattern (continuous outcome, `c1 = 1`, default grid, seeds
  1 to 10), the selection kept at most one mediator: one in 4 of the 10
  samples and none in the other 6. The power-enhanced test could then
  draw on that one mediator only, not on the contrasting structure; it
  rejected in 4 of the 10 samples, and the benchmark in 2.
- Strongly correlated mediators. Correlation among the mediators weakens
  both the selection and the screen. Under the contrasting pattern
  (continuous outcome, `n = 200`, `p = 60`, `c1 = 1`, default grid, seeds
  1 to 10), the screen selected at least one mediator in 6 of 10 samples
  with an autoregressive correlation of 0.9 among neighboring mediators,
  against 10 of 10 at 0.5, and the selected set varied from sample to
  sample (one sample named a null mediator beside an active one at both
  correlations).
- Selected mediators in finite samples. The screen's control of its
  error rate is asymptotic. Under the contrasting pattern with a strong
  signal (continuous outcome, `c1 = 1`, default grid, Bonferroni screen,
  `error_level = 0.05`), the selected set held an inactive mediator in
  0.14 of 400 samples at `n = 150`, `p = 60` and in 0.11 of 200 samples at
  `n = 300`, `p = 500`; the rate was 0.05 under the homogeneous pattern at
  the smaller design and 0.00 with no signal. It depends on the tuning
  grid: 20 values from 0.2 to 0.39 gave 0.00 in the same runs, with lower
  recall (0.13 against 0.48, and 0.25 against 0.47).
- Rare binary outcomes. With about 2 to 3 percent events at `n = 300`
  the penalized logistic fit selects nothing at any grid value; the fit
  is empty and the function says so.
- Overdispersed counts. The count model is Poisson. In null simulations
  with negative binomial counts (size parameter 1) at `n = 200` and
  `p = 100`, 300 replications on a grid of 20 values from 0.05 to 1, its
  size was not inflated, but about a tenth of the fits were empty.
- Missing values. `pe_mediation()` stops on a missing value, so supply
  complete cases; `pe_mediate()` drops incomplete rows and says how many
  it dropped.

## References

Yu, X., & Kelley, K. (in press). Power Enhancement in High-Dimensional
Heterogeneous Mediation Analysis. *Journal of the American Statistical
Association*.

Fan, J., Liao, Y., & Yao, J. (2015). Power enhancement in
high-dimensional cross-sectional tests. *Econometrica, 83*(4),
1497--1541. https://doi.org/10.3982/ECTA12749

Guo, X., Li, R., Liu, J., & Zeng, M. (2022). High-dimensional mediation
analysis for selecting DNA methylation loci mediating childhood trauma
and cortisol stress reactivity. *Journal of the American Statistical
Association, 117*(539), 1110--1121.
https://doi.org/10.1080/01621459.2022.2053136

## Session Information

```{r session-info}
sessionInfo()
```
