---
title: "Reproducing the Article's Analyses With POEMED"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Reproducing the Article's Analyses With POEMED}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.2)
```

This vignette accompanies the article "Power Enhancement in
High-Dimensional Heterogeneous Mediation Analysis" (Yu and Kelley, in
press at the *Journal of the American Statistical Association*). Its
second part reruns the article's empirical data analysis (its Section 5)
with POEMED and reads the package's results against the printed tables,
cell by cell. Its first part describes the article's simulation designs
(its Section 4, and Sections S.4.1 and S.4.4 of its supplement) and gives
the calls that run each of them at the article's scale, on the grids of
the article's own scripts where the article records one. The companion
vignette *Power Enhancement for
High-Dimensional Mediation: A Guided Tour* introduces the method itself.

**Scope of the comparison.** The article benchmarks the power-enhanced
tests (PE-HDMM for linear outcomes, PE-HDGMM for generalized outcomes)
against several competitors: HDMM, HILMA, and GlobalTest. POEMED
implements the proposed PE tests and the **HDMM** benchmark (the
total-indirect-effect Wald test that the PE component augments). Every
POEMED result therefore reports the central contrast of the article's
figures and tables, **PE versus HDMM**; HILMA and GlobalTest are
separate methods outside this package.

Load the package and fix the random-number seed before running any of
the code below:

```{r seed}
library(POEMED)
set.seed(113)
```

# Part I. The Article's Simulation Designs (Article Section 4)

The simulation study asks two questions of every test. Under the global
null of no active mediator (`c1 = 0`) the rejection rate estimates the
**Type I error rate**, which should sit near the nominal level,
`alpha_level = 0.05`. Away from the null (`c1 != 0`) the rejection rate
estimates **power**. `pe_power_curve()` sweeps `c1` and returns both the
benchmark (`rejection_hdmm`) and power-enhanced (`rejection_pe`)
rejection rates at each grid point, and `pe_simulation_study()` runs
several patterns at once and stacks the results.

The article's designs use `n = 300`, `p = 500`, and **1,000
replications** per design point, and its scripts search a tuning grid
set by hand for each outcome family: 20 values from 0.2 to 0.39 for the
linear model, 15 values from 0.04 to 0.2 for the logistic model, and 15
values from 0.7 to 5 for the Poisson model. (The linear script searched a
second grid, 20 values from 0.27 to 0.46, for the reduced model that the
benchmark test needs; the package searches one grid for both fits, so the
calls below pass the full-model grid.) The calls below for Sections 4.1,
4.2, and S.4.1 pass those grids as `lambda_grid`, so they run the
article's designs as its scripts did, whatever the package's default grid
(100 values from 0.05 to 10); the article records no grid for the
real-data-motivated setting of Section S.4.4, whose call uses the package
default. The chunks in this
part are shown but not run: a linear or logistic panel takes one to a
few hours of computing at this scale, and the Poisson homogeneous panel
days. Set `cores` above 1 to parallelize; a seeded run repeats exactly
at the same number of cores.

## Section 4.1: Linear Mediation Models

The article studies a univariate exposure with `n = 300` continuous
outcomes and `p = 500` candidate mediators whose noise has an
autoregressive `0.5^|i-j|` covariance. The mediator-exposure coefficient
is `Gamma_x = c1 * tau`, where `tau` is 0.1, 0.2, 0.3, 0.4, 0.5 for the
first five mediators and an independent N(0, 0.5^2) draw for each of the
others, the direct effect is
`c2 = 0.5`, and two outcome-mediator patterns are contrasted:

* **Setting (i), homogeneous:** `alpha_m = (1, 0.8, 0.6, 0.4, 0.2, 0, ...)`,
  so all active mediators push the same way and the total indirect effect
  is `beta = 0.7 * c1` (nonzero). This is the regime where ordinary tests
  already work; the question is whether PE *adds* power.
* **Setting (ii), contrasting:** `alpha_m = (1, -0.5, 0.4, -0.3, 0, ...)`,
  four active mediators whose effects cancel so the total indirect effect
  is `beta = 0` for every `c1`. This is the regime that defeats a
  total-indirect-effect test, and where PE is designed to rescue power.

POEMED maps these to `pattern = "homogeneous"` and `pattern = "contrasting"`
in `simulate_mediation_data()` and `pe_power_curve()`. The article's
Figure 3 sweeps `c1` from -1 to 1 (the symmetric design fills in the
negative half), one panel per pattern:

```{r lm-paper-scale, eval = FALSE}
c1_grid <- seq(-1, 1, by = 0.1)
lm_study <- pe_simulation_study(n = 300, p = 500, outcome = "continuous",
                                patterns = c("homogeneous", "contrasting"),
                                c1_grid = c1_grid, c2 = 0.5, n_rep = 1000,
                                lambda_grid = seq(0.2, 0.39, length.out = 20),
                                cores = 4, seed = 113)
```

The result has one row per pattern and `c1`, with the two rejection
rates and the number of replications whose penalized fit selected no
mediator (`n_empty`). `plot()` on a `pe_power_curve()` result draws the
two curves against `c1`, as the guided-tour vignette shows.

### The Real-Data-Motivated Heterogeneous Setting (Supplement Section S.4.4)

The supplement adds a third, harder pattern in its Section S.4.4: a
setting calibrated to the DNA-methylation case study of Guo et al.
(2022), with **eleven active mediators of mixed sign** that neither all
agree (homogeneous) nor exactly cancel (contrasting).
`simulate_guo_mediation()` generates it from the calibration constants
shipped as `guo_calibration`; the total indirect
effect is `guo_calibration$beta_per_c1 * c1`, about `-1.597 * c1`. (The
article reports `-1.5977`, computed from the full-precision Guo
coefficients; the package ships those coefficients rounded to the three
decimals printed in the supplement, which give `-1.597`.)

```{r guo-data}
d <- simulate_guo_mediation(c1 = 1, seed = 113)
dim(d$M)                              # 85 x 1008, the case-study dimensions
d$active_mediators                    # the 11 active loci
round(d$beta, 4)                      # beta_per_c1 * c1, i.e. about -1.597
```

The supplement's Figure S.14 sweeps `c1` over `c(0, +/- 0.1, ..., +/- 1)`
at the case-study size `n = 85`, without the calibrated confounders in
its panel (a) and with them in its panel (b), averaged over 1,000
replications. The article records no tuning grid for this setting, so
the fits below use the package default (the example on
`?simulate_guo_mediation` passes a shorter grid to stay fast and selects
the same loci as the default):

```{r guo-paper-scale, eval = FALSE}
rej <- function(c1, confounders) {
  ps <- replicate(1000, {
    d <- simulate_guo_mediation(c1 = c1, confounders = confounders)
    f <- pe_mediation(d$X, d$Y, d$M, Z = d$Z, outcome = "continuous")
    c(f$value[f$term == "pval_hdmm"], f$value[f$term == "pval_pe"])
  })
  rowMeans(ps <= 0.05)
}
guo_without <- sapply(seq(-1, 1, by = 0.1), rej, confounders = FALSE)
guo_with <- sapply(seq(-1, 1, by = 0.1), rej, confounders = TRUE)
```

## Section 4.2: Mediation Models With Generalized Linear Outcomes

The same construction is repeated for non-continuous outcomes through a
link function: a **logistic** model for binary outcomes (PE-HDGMM versus
HDGMM) and a **Poisson** model for counts. Again `n = 300`, `p = 500`,
with a homogeneous and a contrasting pattern apiece; the article fixes
the direct effect at `c2 = 1` for the logistic models and `c2 = 0.4` for
the Poisson models. The calls below run the four panels of the article's
Figures 4 and 5. The Poisson homogeneous fits take tens of seconds each
at this design, so that panel alone needs well over 100 hours of
computing (days, even on 4 cores); each of the other three panels needs
a few hours.

```{r glm-paper-scale, eval = FALSE}
c1_grid <- seq(-1, 1, by = 0.1)
# Logistic: c2 = 1
glm_logit <- pe_simulation_study(n = 300, p = 500, outcome = "binary",
                                 patterns = c("homogeneous", "contrasting"),
                                 c1_grid = c1_grid, c2 = 1, n_rep = 1000,
                                 lambda_grid = seq(0.04, 0.2, length.out = 15),
                                 cores = 4, seed = 113)
# Poisson: c2 = 0.4
glm_pois <- pe_simulation_study(n = 300, p = 500, outcome = "count",
                                patterns = c("homogeneous", "contrasting"),
                                c1_grid = c1_grid, c2 = 0.4, n_rep = 1000,
                                lambda_grid = seq(0.7, 5, length.out = 15),
                                cores = 4, seed = 113)
```

## Supplement Section S.4.1: Identifying Individual Mediators (FWER and FDR)

Beyond the global test, the article's supplement studies how well the
PE machinery **identifies which** individual mediators are active, and
whether it controls the familywise error rate (FWER) or false discovery
rate (FDR). POEMED exposes this through the `method` argument
(`"Bonferroni"` for FWER, `"BH"` or `"BY"` for FDR) and the
`report_all_methods` switch, which reports all three side by side as the
supplement does; for a single fit, `pe_selection()` shows the selected
set and the power-enhanced p-value under each method (the guided-tour
vignette shows one).

`pe_identification_study()` runs the supplement's identification
*study*: it scores each method's selections against the truth set across
a grid of `c1` and reports the empirical FWER, FDR, precision, and
recall, with the supplement's definitions. A mediator counts as truly
active when its outcome coefficient is nonzero, at every `c1` including
`c1 = 0` (five mediators under the homogeneous pattern, four under the
contrasting pattern); pass `truth = "mediation"` for the both-paths
definition, under which nothing is active at `c1 = 0` and `recall` is
`NA` there. The supplement's Tables S.1 to S.3 come from the calls
below, one per pattern:

```{r id-study-paper, eval = FALSE}
id_study <- lapply(c("homogeneous", "contrasting"), function(pattern)
  pe_identification_study(n = 300, p = 500, outcome = "continuous",
                          pattern = pattern, c1_grid = seq(0, 1, by = 0.1),
                          n_rep = 1000,
                          lambda_grid = seq(0.2, 0.39, length.out = 20),
                          cores = 4, seed = 113))
```

# Part II. Empirical Data Analysis (Article Section 5)

The article asks whether health-care expenditure mediates the
relationship between economic growth and population health across the
world's countries. POEMED ships that exact data set, `WHO_health_mediation`:
a country-by-year panel (2000--2021) with

* **Exposure** `X`: annual growth in GDP per capita (World Bank),
* **Mediators** `M`: 57 health-expenditure indicators (WHO Global Health
  Expenditure Database),
* **Outcomes** `Y`: five health outcomes, infant mortality (`imr`),
  under-five mortality (`u5mr`), life expectancy (`leb`), low birthweight
  (`lbw`), and undernourishment (`pou`),

with WHO region and World Bank income group available as the grouping and
confounding variables.

```{r who-data}
data(WHO_health_mediation)
dim(WHO_health_mediation)
table(WHO_health_mediation$region)
```

## Section 5, Table 1: Infant Mortality (IMR)

`WHO_mediation_analysis()` builds a table in the layout of the article's
per-outcome tables: it fits the global model (region, income, and year
as confounders), then a separate model within each WHO region (income
and year as confounders) and each income group (region and year as
confounders), and reports the HDMM and PE p-values for the total
indirect effect together with the mediators the PE component selects.

The function searches one tuning grid, `seq(0.1, 2, length.out = 100)`,
for the global all-country model and for every subgroup model alike. It
is the grid behind the article's real-data tables (the article's text
gives no grid; it states only that the tuning parameter is chosen by the
HBIC criterion), and it is the default of both `lambda_grid`, for the
subgroup models, and `lambda_grid_global`, for the global model, so no
explicit grid arguments are needed.

```{r who-imr}
imr <- WHO_mediation_analysis("imr", groupings = c("global", "region", "income"))
imr[, c("group", "n_countries", "pval_hdmm", "pval_pe", "selected_mediators")]
```

```{r who-imr-values, include = FALSE}
p_hdmm <- function(g) formatC(signif(imr$pval_hdmm[imr$group == g], 2),
                              digits = 2, format = "fg", flag = "#")
p4 <- function(g) sprintf("%.4f", imr$pval_hdmm[imr$group == g])
```

Read this against the article's Table 1. The highlights:

* **Global (ALL):** general government expenditure as a share of GDP
  (`gge_gdp`) is the mediator the screen selects; HDMM is borderline
  (`pval_hdmm` = `r p_hdmm("ALL")`) while PE is essentially zero.
* **SEAR (South-East Asia):** HDMM sees nothing (`pval_hdmm` =
  `r p_hdmm("SEAR")`) while PE detects external health expenditure per
  capita in constant 2021 US dollars (`ext_usd2021_pc`).
* **WPR (Western Pacific):** HDMM sees nothing (`pval_hdmm` =
  `r p_hdmm("WPR")`) while PE detects compulsory health insurance as a
  share of current health expenditure (`chi_che`).
* **Low income:** HDMM sees nothing (`pval_hdmm` = `r p_hdmm("Low")`)
  while PE detects domestic private health expenditure (`pvtd_usd2021`).
* **EMR (Eastern Mediterranean), EUR (Europe), and high income:** the
  benchmark rejects (`pval_hdmm` = `r p_hdmm("EMR")`, `r p_hdmm("EUR")`,
  and `r p_hdmm("High")`; Table 1 prints the EUR value as 0.0001). In the
  high-income group PE identifies out-of-pocket spending and social
  health insurance as shares of current health expenditure (`oops_che`,
  `shi_che`); in EMR and EUR no single indicator passes the screen, so
  the PE p-value equals the benchmark's.
* **AFR, AMR, lower-middle income, and upper-middle income:** neither
  test detects mediation.

In the global model, SEAR, WPR, and the low-income group, heterogeneous
mediation hides the signal from the conventional test and the PE
component recovers it. Among the income groups the strongest signals
appear in the low- and high-income groups, as Yu and Kelley (in press)
report in Table 1.

The table above agrees with the article's Table 1 row for row. Every
group's identified set is the one the table lists, and every benchmark
p-value matches the printed value to four decimals except SEAR, which is
`r p4("SEAR")` here where the table prints 0.6482. Where the table
prints a PE p-value below `1e-12`, the package prints `< 0.0001`, its
display floor; the values behind those entries are below `1e-12` as
well.

`summary()` digests the whole table, highlighting the groups where the
power-enhanced test detects mediation the benchmark misses (the article's
Table 1 sets those four rows in bold):

```{r who-summary}
summary(imr)
```

The supplement also gives an **extended** version of each table (its
Table S.8 for IMR) that breaks the power-enhanced result out by
multiplicity method. Pass
`full_table = TRUE` to get that layout, with the PE p-value and selected
mediators reported separately under Bonferroni (FWER), BH, and BY (FDR):

```{r who-full}
WHO_mediation_analysis("imr", groupings = "global", full_table = TRUE)
```

## The Other Outcomes, and the Region-by-Income Cells

The remaining article outcomes (undernourishment in the article's
Table 2, and life expectancy, under-five mortality, and low birthweight
in the supplement's Tables S.10 to S.12) run identically; only the
outcome code changes:

```{r who-others, eval = FALSE}
for (y in c("u5mr", "leb", "lbw", "pou"))
  print(WHO_mediation_analysis(y, groupings = c("global", "region", "income")))
```

The article also drills into the **region-by-income** cells.
`WHO_mediation_analysis()` returns the region-by-income block of the
table in one call, with `NA` for the cells that have no countries or
whose fit fails:

```{r who-region-income, eval = FALSE}
imr_cells <- WHO_mediation_analysis("imr", groupings = "region_income")
```

To look inside one cell, build it with `WHO_mediation_design()`, which
assembles the exposure, outcome, mediator, and confounder blocks for a
chosen subset, and fit it the way `WHO_mediation_analysis()` does. A
small cell can contain indicators that never vary inside it. Such a
column carries no signal and cannot be standardized, so the code below
drops it by hand before standardizing, as the function does, and the
print footer then names the selected mediators by their indicator codes.
The code also passes the function's default grid, so this fit is the
AFR-Upper-middle row of the block above:

```{r who-cell}
cell <- WHO_mediation_design("imr", region = "AFR", income = "Upper-middle")
c(n_countries = cell$n_countries, mediators = ncol(cell$M))

keep <- apply(cell$M, 2L, sd) > 0          # FALSE for a constant indicator
cell$indicators[!keep]
fit <- pe_mediation(scale(cell$X), cell$Y - mean(cell$Y),
                    scale(cell$M[, keep, drop = FALSE]),
                    Z = cell$Z, outcome = "continuous", scale = FALSE,
                    lambda_grid = seq(0.1, 2, length.out = 100))
fit[fit$term %in% c("pval_hdmm", "pval_pe"), ]
```

The selected mediator, domestic private health expenditure per capita in
constant 2021 national currency units (`pvtd_ncu2021_pc`), is the one the
article's Table 1 lists for this cell, and the benchmark p-value,
`r sprintf("%.4f", fit$value[fit$term == "pval_hdmm"])`, is the value the
table prints.

## Six Cells Reported as NA

Across the region-by-income blocks of the article's tables for its five
outcomes (Table 1, Table 2, and Tables S.8 to S.12), six cells that the
article prints with a value come back from `WHO_mediation_analysis()` as
`NA`, each with a warning that names the cell and quotes the reason.
They are of two kinds.

**Four rank-deficient cells.** In the EMR-Lower-middle cells of IMR and
U5MR and the AFR-Upper-middle cells of LEB and U5MR, two of the 57
indicators are exactly collinear: domestic and external health
expenditure as shares of current health expenditure (`dom_che`,
`ext_che`) sum to 100 in every country-year of those cells. The
AFR-Upper-middle cells add a second identity, since compulsory and
social health insurance as shares of current health expenditure
(`chi_che`, `shi_che`) coincide there (compulsory private insurance,
`chi_pvt_che`, is zero throughout, which is why it was dropped as
constant above). The benchmark test inverts the covariance matrix of
the mediators the penalized fit selected, and when that selection keeps
both members of a pair the matrix is singular; the fit stops with the
error that the system is computationally singular, and the function
reports the cell as `NA`. (The same two identities hold in the IMR cell
above, where the selection keeps `ext_che` and `chi_che` without their
partners, so the test goes through.) In three of the four hand fits
below HBIC chooses 0.1, the smallest value of the article's grid, for the
full model, the reduced model, or both; the fits keep that grid as the
article's analysis did, so the warning a single `pe_mediation()` fit
gives at an end of its grid is not shown in this section. The article's
values come back when the collinearity is removed before the fit, as the
article's analysis did: it dropped `dom_che` and `vfa_che` (near-exact partners of `ext_che`
and `cfa_che`) in the EMR-Lower-middle cells, and `shi_che` and `ext_che`
in the AFR-Upper-middle cells:

```{r who-rank-deficient, warning = FALSE}
cell <- WHO_mediation_design("imr", region = "EMR", income = "Lower-middle")
range(cell$M[, "dom_che"] + cell$M[, "ext_che"])   # 100 in every row

fit_cell <- function(outcome, region, income, drop) {
  cell <- WHO_mediation_design(outcome, region = region, income = income)
  keep <- apply(cell$M, 2L, sd) > 0 & !(cell$indicators %in% drop)
  fit <- pe_mediation(scale(cell$X), cell$Y - mean(cell$Y),
                      scale(cell$M[, keep, drop = FALSE]),
                      Z = cell$Z, outcome = "continuous", scale = FALSE,
                      lambda_grid = seq(0.1, 2, length.out = 100))
  selected <- cell$indicators[keep][attr(fit, "selected_mediators")]
  data.frame(outcome = outcome, cell = paste(region, income, sep = "-"),
             dropped = paste(drop, collapse = ", "),
             pval_hdmm = fit$value[fit$term == "pval_hdmm"],
             selected = if (length(selected)) paste(selected, collapse = ", ")
                        else "none")
}
rd <- rbind(fit_cell("imr",  "EMR", "Lower-middle",
                     drop = c("dom_che", "vfa_che")),
            fit_cell("u5mr", "EMR", "Lower-middle",
                     drop = c("dom_che", "vfa_che")),
            fit_cell("leb",  "AFR", "Upper-middle",
                     drop = c("shi_che", "ext_che")),
            fit_cell("u5mr", "AFR", "Upper-middle",
                     drop = c("shi_che", "ext_che")))
rd
```

The four benchmark p-values,
`r paste(sprintf("%.4f", rd$pval_hdmm), collapse = ", ")`, are the values
the article prints for these cells (Table 1 for IMR, Table S.10 for LEB,
and Table S.11 for U5MR), and the U5MR AFR-Upper-middle cell recovers the
mediator Table S.11 lists, `pvtd_ncu2021_pc`.

**Two constant-outcome cells.** In the EUR-High and WPR-High cells of
the undernourishment analysis (PoU, the article's Table 2), the outcome
is 2.5 in every country-year of the cell: the World Bank publishes a
prevalence of undernourishment below 2.5 percent as 2.5, its lower
reporting bound, and the high-income countries of those two regions are
all at that bound. There is no variation in the outcome to mediate, so
`pe_mediation()` refuses the fit with the message that `Y` is constant,
and the function reports the two cells as `NA`. Table 2 prints a p-value
of 1 for both.

```{r who-constant-y}
sapply(c("EUR", "WPR"), function(r)
  unique(WHO_mediation_design("pou", region = r, income = "High")$Y))
```

## Why a Mediator? Opening Up the Screening Step

For any fit, `pe_mediators()` shows the evidence the PE component acted
on: the two path statistics, the screening p-value (the larger of the
two path p-values), and the `selected` flag, for every candidate the
penalized fit selected. This is the per-mediator detail behind the last
column of the article's tables. In this global model HBIC picks 0.1, the
smallest value of the grid, as it does in about half of the article's
cells; the warning a single fit gives at an end of its grid is not shown
here either.

```{r who-screen, warning = FALSE}
des <- WHO_mediation_design("imr")                  # global IMR design
gfit <- pe_mediation(scale(des$X), des$Y - mean(des$Y), scale(des$M),
                     Z = des$Z, outcome = "continuous", scale = FALSE,
                     lambda_grid = seq(0.1, 2, length.out = 100))
pe_mediators(gfit)
```

# Reproducibility Notes

* The global, regional, and income analyses above are deterministic.
  Called with no explicit grid arguments, `WHO_mediation_analysis()`
  searches `seq(0.1, 2, length.out = 100)`, the grid behind the
  article's real-data tables, for the all-country model and for every
  subgroup model, and returns the selected mediators and p-values shown
  here.
* Part I runs no simulation study. Its `eval = FALSE` blocks give the
  calls that run the article's designs at their scale and with 1,000
  replications: `n = 300` and `p = 500` on the grids of the article's
  scripts (passed as `lambda_grid`) for Sections 4.1, 4.2, and S.4.1,
  and the case-study `n = 85` and `p = 1008` on the package default grid
  for supplement Section S.4.4, for which the article records no grid.
  Set `cores` above 1 to parallelize and
  `seed` for exact reproducibility; a seeded run with `cores` above 1
  repeats exactly at the same number of cores.
* Exact real-data p-values depend on the preprocessing. The article's
  text says that covariates are standardized before fitting.
  `WHO_mediation_analysis()` instead applies the preprocessing of the
  authors' analysis code behind the article's tables: it standardizes the
  exposure and mediators, centers the outcome, and leaves the confounders
  (the region and income indicators and the year) unscaled. The region
  and income factors enter through treatment contrasts with alphabetical
  reference levels. The guided-tour vignette shows how to do this by hand
  and what standardizing the confounders as well changes.

# References

Yu, X., & Kelley, K. (in press). Power Enhancement in High-Dimensional
Heterogeneous Mediation Analysis. *Journal of the American Statistical
Association*.

Guo, X., Li, R., Liu, J., & Zeng, M. (2022). High-dimensional mediation
analysis for selecting DNA methylation loci mediating childhood trauma
and cortisol stress reactivity. *Journal of the American Statistical
Association, 117*(539), 1110--1121.
https://doi.org/10.1080/01621459.2022.2053136

# Session Information

```{r session-info}
sessionInfo()
```
